### 2024-05-13 thesis/chapters/03-design.tex
Thesis send to feedback

\chapter{Design} \label{chapter:design}
During the design stage, we set out research questions and detail the steps of the machine learning pipeline for diagnosing fault types in the MaFaulDa dataset. The explanatory analysis of MaFaulDa hints at the underlying dependencies with feature sets that can degrade the machine learning model of choice. We adapt the methodology for measurements from standards and apply it to data logger requirements and data acquisition from industrial equipment.

\section{Research questions}
The thesis aims to provide answers to four research questions. The focus is primarily on making data flow more efficient in an industrial sensor network that monitors rotating machines. The \textbf{research questions} are:
\begin{enumerate}[label=RQ\arabic*., font=\bfseries]
    \itemsep0pt
	\item Which temporal and spectral features can be extracted from vibration signals to provide the most accurate record of machinery faults?
	\item What is the reduction in transmission goodput when chosen signal features are used?
	\item What accuracy of prediction models can be achieved with various feature subsets?
	\item How can machinery faults be continuously identified and predicted based upon collected events?
\end{enumerate}

\noindent In accomplishing the objectives of our research we propose several \textbf{goals}:
\begin{todolist}
    \itemsep0pt
    \item Statistically and visually describe vibration signals from the Machinery fault database (MauFaulDa).
    \item Establish a list of conditions that should be later investigated in the experimental setting.
    \item Prepare dataset to be used in conjunction with machine learning models, namely by identifying labels and balancing classes.
    \item Find the best subsets of features in the time and frequnecy domain with previously analyzed feature extraction and selection methods.
    \item Evaluate the performance of models described in the diagnostics section with a significant focus on the k-nearest neighbor algorithm.
    \item Acquire measurements of vibrations from machines in the real environment to form a novel dataset of machinery behavior.
    \item Develop hardware and implement its firmware to obtain such measurements of the quality demanded by vibrodiagnostics standards.
\end{todolist}

@@

We neglect the acceleration sensor orientation, therefore, the labels for shaft misalignment in vertical and horizontal directions are merged into the same group. In the end, that leaves \textbf{6 types of labels}: baseline, two shaft faults are imbalance and misalignment, and three bearing faults are cage fault, ball fault, and outer race fault.

Depending on the chosen bearing position, only time series associated with that bearing are retained because the standards demand each bearing be assessed separately. Fault classification concerns bearing in direct contact and shaft mechanically passing through it.

@@

The final \textbf{file description} is the triplet of \textbf{fault}, \textbf{severity}, and \textbf{rpm}.

After labeling we insert a step to split the time series into parts when the duration is long enough to introduce more observations artificially. This assumption is not met for MaFaulDa because, for the desired frequency resolution for features of close to 1~Hz, we need a window size of 32768 samples. Variation reduction in spectral estimation brings the necessity to average at least 12 overlapping windows. The full length of 5 seconds provides 15 whole overlapping windows.

The DC component in the three-dimensional vibration signal is removed by subtracting the global mean. Immediately follows a digital IIR Butterworth low pass filter of \nth{5} order with cutoff frequency 10 kHz at -3 dB. Before the usage of a low pass filter, the peak at 20 kHz with a sideband was present as an unwanted artifact. It could not have been reliably recorded due to the linear frequency response of the sensor up to 10 kHz. At the same time, such a frequency is outside the range of any feasible MEMS accelerometer.

Initial labels can be exchanged for a label of fault-free state when the severity is low. The count of severity levels is not identical in every group. The numerical amounts for each level are sorted in ascending order within each fault and then these levels are normalized using a min-max scaler.

The preprocessed signal from each axis of accelerometer is packed up to two base feature sets in time domain (TD) and frequency domain (FD). The feature extraction formulas are is identical with formulas presented in the section \ref{section:feature-extraction} of analysis chapter. The frequency transform is Welch's method averaging FFT vectors from 32768 samples after Hann windowing and 50\% overlap. The feature sets are:
\begin{itemize}
\itemsep0pt
\item \textbf{TD has 10 time-domain features} - peak-to-peak amplitude (\emph{pp}), zero-crossing rate (\emph{zerocross}), root mean square (\emph{rms}), skewness, kurtosis, shape factor, crest factor, impulse factor, clearance factor, average amplitude change (aac), 
\item \textbf{FD has 11 frequency-domain features} - spectral centroid, standard deviation (\emph{std}), skewness, kurtosis, roll-on frequency, roll-off frequency, spectral flux, noisiness, spectral negentropy, energy, entropy.
\end{itemize}

The four conditions are applied in the instance of the pipeline to filter observations with calculated features to create 24 scenarios:
\begin{itemize}
\itemsep0pt
\item \textbf{Source bearings} - samples and fault labels are left in either just for the inner bearing (A), for only the outer bearing (B), or for both bearings (A+B).
\item \textbf{Feature domain} - the set of extracted features as an input to downstream models is changed either to TD or FD set,
\item \textbf{Accelerometer axis} - Either one principal direction of motion or all three dimensions is aggregated under same feature name (1 or 3),
\item \textbf{High severity faults} - to more precisely simulate fault frequencies found in real environment the low severity defects can be considered as normal operation (Yes or No).
\end{itemize}

The fault classification performance metrics are evaluated after balancing classes feature normalization, and 5-fold cross-validation validation on the k-nearest neighbor classifier with Euclidean distance metric. The population of classes is evened out with random oversampling using a strategy to resample all but the majority class. In the multi-class classification, the micro-averaged performance metrics of accuracy, precision, and recall result in the same number. Therefore, accuracy is only one on the output.

The hyperparameter of k-neighbours and a subset of features are tweaked, and results are compared under different preprocessing conditions and in multiple experiments:

@@

Out of the 1951 time series, the inner bearing (A) has 1438, and the outer bearing (B) has 1393~(Tab.~\ref{tab:design:label-count}). Sorce bearing A+B just concatanes observations from both postions so the column sums their counts together. The observations for shaft defects are shared regardless of the bearing. 
\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|ll|r|r|r|r|r|r|}
\hline
\multicolumn{2}{|l|}{\textbf{Source bearing}}                                            & \multicolumn{1}{l|}{A}  & \multicolumn{1}{l|}{A+B} & \multicolumn{1}{l|}{B}  & \multicolumn{1}{l|}{A}   & \multicolumn{1}{l|}{A+B} & \multicolumn{1}{l|}{B}   \\ \hline
\multicolumn{2}{|l|}{\textbf{High severity faults}}                                & \multicolumn{1}{l|}{No} & \multicolumn{1}{l|}{No}  & \multicolumn{1}{l|}{No} & \multicolumn{1}{l|}{Yes} & \multicolumn{1}{l|}{Yes} & \multicolumn{1}{l|}{Yes} \\ \hline
\multicolumn{1}{|l|}{\multirow{6}{*}{\textbf{Fault}}} & \textbf{misalignment}     & 498                     & 996                      & 498                     & 248                      & 496                      & 248                      \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{imbalance}        & 333                     & 666                      & 333                     & 188                      & 376                      & 188                      \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{cage fault}       & 188                     & 376                      & 188                     & 91                       & 181                      & 90                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{ball fault}       & 186                     & 323                      & 137                     & 87                       & 132                      & 45                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{outer race fault} & 184                     & 372                      & 188                     & 86                       & 176                      & 90                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{normal}           & 49                      & 98                       & 49                      & 738                      & 1470                     & 732                      \\ \hline
\multicolumn{2}{|l|}{$\Sigma$}                                                    & \textbf{1438}           & \textbf{2831}            & \textbf{1393}           & \textbf{1438}            & \textbf{2831}            & \textbf{1393}            \\ \hline
\end{tabular}
\caption{Number of observation in MaFaulDa split by class label according to source bearing}
\label{tab:design:label-count}
\end{table}

>>>>

\chapter{Design} \label{chapter:design}
During the design stage, we set out research questions and detail the steps of the machine learning pipeline for diagnosing fault types in the MaFaulDa dataset. The explanatory analysis of MaFaulDa hints at the underlying dependencies within feature sets that can degrade the machine learning model of choice. We apply the methodology for vibration measurements from standards to the data logger requirements and data acquisition from industrial equipment.

\section{Research questions}
The thesis aims to provide answers to five research questions. The main focus is on making data flow more efficient in an industrial sensor network that monitors rotating machines. The \textbf{research questions} are:
\begin{enumerate}[label=RQ\arabic*., font=\bfseries]
    \itemsep0pt
	\item Which time-domain and frequency-domain features are extractable from vibrations to provide the most accurate record of machinery faults?
	\item What is the reduction in transmission goodput with signal features?
	\item What accuracies of fault classification are achievable with tiny feature subsets?
	\item How do the vibration signals collected on the industrial machinery behave?
	\item How can machinery faults be identified over time from collected events?
\end{enumerate}

\noindent In accomplishing the objectives of our research, we propose several \textbf{goals}:
\begin{todolist}
    \itemsep0pt
    \item Statistically and visually describe vibration signals from the Machinery fault database (MauFaulDa).
    \item Establish a list of conditions that will be investigated later in the experimental setting.
    \item Prepare the dataset to be used in conjunction with machine learning models namely by identifying labels and balancing classes.
    \item Find the best subsets of features in the time and frequency domain with previously analyzed feature extraction and selection methods.
    \item Evaluate the performance of models described in the diagnostics section with a significant focus on the k-nearest neighbor algorithm.
    \item Acquire measurements of vibrations from machines in the real environment to form a novel dataset of machinery behavior.
    \item Develop hardware and implement its firmware to obtain such measurements of the quality demanded by vibrodiagnostics standards.
\end{todolist}

@@

We neglect the \textbf{acceleration sensor orientation}, therefore, the labels for shaft misalignment in vertical and horizontal directions are merged into the same group. In the end, that leaves \textbf{6 types of labels}: baseline, two shaft faults are imbalance and misalignment, and three bearing faults are cage fault, ball fault, and outer race fault.

Depending on the \textbf{chosen bearing position}, only time series associated with that bearing are retained because the standards demand each bearing be assessed separately. Fault classification concerns bearing in direct contact and shaft mechanically passing through it.

@@

\textbf{File description} is the triplet of \textbf{fault}, \textbf{severity}, and \textbf{rpm}.

After labeling we insert a step to split the time series into parts when the duration is long enough to introduce more observations artificially. This assumption for splitting is not met for MaFaulDa because, for the desired frequency resolution for features of close to 1~Hz, we need a window size of 32768 samples. Variation reduction in spectral estimation brings the necessity to average at least 12 overlapping windows. The full length of 5 seconds provides 15 whole overlapping windows.

The \textbf{DC component} in the three-dimensional vibration signal is removed by subtracting the global mean. Immediately follows a digital IIR Butterworth low pass filter of \nth{5} order with cutoff frequency 10 kHz at -3 dB. Before the usage of a low pass filter, the peak at 20 kHz with a sideband was present as an unwanted artifact. It could not have been reliably recorded due to the linear frequency response of the sensor up to 10 kHz. At the same time, such a frequency is outside the range of any feasible MEMS accelerometer.

\textbf{Initial labels can be exchanged} for a label of fault-free state when the severity is low. The count of severity levels is not identical in every group. The numerical amounts for each level are sorted in ascending order within each fault and then these levels are normalized using a min-max scaler.

The preprocessed signal from each axis of accelerometer is packed up to \textbf{two base feature sets} in time domain (TD) and frequency domain (FD). The feature extraction formulas are is identical with formulas presented in the section \ref{section:feature-extraction} of analysis chapter. The frequency transform is Welch's method averaging FFT vectors from 32768 samples after Hann windowing and 50\% overlap. The feature sets are:
\begin{itemize}
\itemsep0pt
\item \textbf{TD has 10 time-domain features} - peak-to-peak amplitude (\emph{pp}), zero-crossing rate (\emph{zerocross}), root mean square (\emph{rms}), skewness, kurtosis, shape factor, crest factor, impulse factor, clearance factor, average amplitude change (aac), 
\item \textbf{FD has 11 frequency-domain features} - spectral centroid, standard deviation (\emph{std}), skewness, kurtosis, roll-on frequency, roll-off frequency, spectral flux, noisiness, spectral negentropy, energy, entropy.
\end{itemize}

The four conditions are applied in the instance of the pipeline to filter observations with calculated features to create \textbf{24 scenarios}:
\begin{itemize}
\itemsep0pt
\item \textbf{Source bearings} - samples and fault labels are left in either just for the inner bearing (A), for only the outer bearing (B), or for both bearings (A+B).
\item \textbf{Feature domain} - the set of extracted features as an input to downstream models is changed either to TD or FD set,
\item \textbf{Accelerometer axis} - Either one principal direction of motion or all three dimensions is aggregated under same feature name (1 or 3),
\item \textbf{High severity faults} - to more precisely simulate fault frequencies found in real environment the low severity defects can be considered as normal operation (Yes or No).
\end{itemize}

The \textbf{fault classification} performance metrics are evaluated after balancing classes feature normalization, and 5-fold cross-validation validation on the k-nearest neighbor classifier with Euclidean distance metric. The population of classes is evened out with random oversampling using a strategy to resample all but the majority class. In the multi-class classification, the micro-averaged performance metrics of accuracy, precision, and recall result in the same number. Therefore, accuracy is only one on the output.

The hyperparameter of k-neighbours and a feature subset size are tweaked, and results are compared under different preprocessing conditions and in multiple experiments:

@@

Out of the 1951 time series, the inner bearing (A) has 1438, and the outer bearing (B) has 1393~(Tab.~\ref{tab:design:label-count}). Source bearing A+B just concatanes observations from both postions so the column sums their counts together. The observations for shaft defects are shared regardless of the bearing. 
\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|ll|r|r|r|r|r|r|}
\hline
\multicolumn{2}{|l|}{\textbf{Source bearing}}                                            & \multicolumn{1}{l|}{A}  & \multicolumn{1}{l|}{A+B} & \multicolumn{1}{l|}{B}  & \multicolumn{1}{l|}{A}   & \multicolumn{1}{l|}{A+B} & \multicolumn{1}{l|}{B}   \\ \hline
\multicolumn{2}{|l|}{\textbf{High severity faults}}                                & \multicolumn{1}{l|}{No} & \multicolumn{1}{l|}{No}  & \multicolumn{1}{l|}{No} & \multicolumn{1}{l|}{Yes} & \multicolumn{1}{l|}{Yes} & \multicolumn{1}{l|}{Yes} \\ \hline
\multicolumn{1}{|l|}{\multirow{6}{*}{\textbf{Fault}}} & \textbf{misalignment}     & 498                     & 996                      & 498                     & 248                      & 496                      & 248                      \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{imbalance}        & 333                     & 666                      & 333                     & 188                      & 376                      & 188                      \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{cage fault}       & 188                     & 376                      & 188                     & 91                       & 181                      & 90                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{ball fault}       & 186                     & 323                      & 137                     & 87                       & 132                      & 45                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{outer race fault} & 184                     & 372                      & 188                     & 86                       & 176                      & 90                       \\ \cline{2-8} 
\multicolumn{1}{|l|}{}                                & \textbf{normal}           & 49                      & 98                       & 49                      & 738                      & 1470                     & 732                      \\ \hline
\multicolumn{2}{|l|}{$\Sigma$}                                                    & \textbf{1438}           & \textbf{2831}            & \textbf{1393}           & \textbf{1438}            & \textbf{2831}            & \textbf{1393}            \\ \hline
\end{tabular}
\caption{Number of observation in MaFaulDa split by class label according to source bearing}
\label{tab:design:label-count}
\end{table}

### 2024-05-13 thesis/chapters/05-evaluation.tex
Thesis send to feedback

\chapter{Evaluation} \label{chapter:evaluation}
The verification of proposed solutions for machinery fault diagnostics is focused on two complementary activities, these are vibration measurement and defect identification. The accuracy of supervised learning using the k-nearest neighbour classifier is determined by the MaFaulDa dataset under various experiments. Data logger recording is compared to the known reference, and the collected Pump dataset is analyzed.

\section{MaFaulDa in k-nearest neighbors}
The four outlined experiments on MaFaulDa involve testing sets of features with very few members to see the effects in predictions when reducing the information content about machine status to the minimum. First, the attributes are left in the complete count after extraction. Then, all of their combinations are enumerated, and the resulting accuracy distribution is matched against accuracy after feature selection with similarity metrics.

@@

The label ``normal'' is not falsely attributed to other classes in either feature set, but other classes can get assigned to be ``normal''. Most mistakes happen while predicting outer race fault, which is confused with cage fault, imbalance, and less frequency with misalignment. The shaft imbalance is used for simulating bearing faults, which is a natural reason for this high error rate.

@@

Under every circumstance, the magnitude of the triaxial feature vector reaches better accuracy than those from only the axis of motion for the same source domain and bearing. The model for inner bearing A is more accurate than outer bearing B. The TD set is generally better in predictions than the FD set for the equivalent k-value. The relabeled dataset for high severity has a steeper decrease in accuracy for the same value of neighbours.

\subsection{Feature subset combinations}
The complete sets of predictors are even greatly shrunken to representation that could be presented in 3D plot or in perpendicular cross sections. These trend variables could be used in distingushing faults the same way as rms amplitude indicates their presence. Each possible combinations of pairs, triplets, and quadruplets constructs a separate k-NN model on which prediction accuracy is evaluated.

@@

The decrease in accuracy with additional neighbours is apparent and similar to the trend in complete sets of features. We are interested in maximum accuracy in the distribution because that is the optimal model feature selection method to try to approach. Its reduction is more noticeable between three and five features 3 - 5\%, and almost the same amount between five and eleven neighbours.

@@

Table~\ref{tab:evaluation:fsel} compares the concrete case of choosing the three features from each domain on bearing A and k-NN with five neighbours. Table~\ref{tab:evaluation:fsel-severity} uses labels for high-severity faults. There is a significant difference in train and test accuracy, which means the model is likely overfitting. The percentile within the distribution is measured to its respective data, the train or validation set. However, the percentile of best features in the test is calculated against train distribution.

Combining the rankings from several metrics is necessary to get consistent results. This is evidenced by the variability of the success of selected feature sets in final prediction performance under multiple conditions. The PCA with 3 components out of complete feature sets is comparable in accuracy to selection methods with original attributes.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-TD-train.png}
        \caption{Time-domain features (train)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-FD-train.png}
        \caption{Frequency-domain features (train)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-TD-test.png}
        \caption{Time-domain features (test)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-FD-test.png}
        \caption{Frequency-domain features (test)}
    \end{subfigure}
    \caption{Model accuracy statistical distibutions with feature selection methods for three predictors (k = 5)}
    \label{fig:evaluation:fsel-model-distr}
\end{figure}

@@

Under 216 scenarios, the mutual information has better median accuracy (80.87\%) and distribution percentile (91.81\%) followed by rank product with an accuracy of 79.82\% of and percentile of 88.97\% (Tab.~\ref{tab:evaluation:compare-fsel-accuracy}). The scenarios are composed of 24 base dataset modifications and options for hyperparameters k-value and number of features.

The rank product was the best method in the majority of 43.52\% scenarios. Mutual information comes second in 40.28\% cases, where it is deemed the best strategy (Tab.~\ref{tab:evaluation:best-selection-method}). The median accuracy in selected cases is also better for rank product 92.38\% compared to 91.79\% mutual information.

@@

\subsection{Incremental learning}
Online learning imitates hardened conditions for machinery diagnostics that appear in deployment. Delayed provision or omission of actual labels undoubtedly degrades the reliability of the classification. The question is how quickly the accuracy approaches the optimal one from the nearest neighbors trained in batch and what is the effect on the classifier from routine difficulties associated with the continous labeling process.

@@

The k-NN models in incremental learning experiments learn on the same base training dataset as in batch learning for bearing A with complete sets of extracted features. In this manner, we can compare the training accuracies in the last sample for both models. Online learning metrics are evaluated by progressive valuation on a dataset that is left unbalanced.

@@

A significant change in the data stream occurs after 294 out of 1438 observations (or after 737 for high severity faults) when all 49 (or 737) normal conditions are consumed in the training process. Counters of other faults show that predictions are skewed towards more represented imbalance and misalignement. The uneven evolution of categories in a stream impacts the development of accuracy in the remaining experiments. The test accuracies of comparable batch models from the three best features are 85.47\% (TD), 87.52\% (FD), 91.71\% (TD severity), and 91.94\% (FD severity).

@@

\section{Industrial dataset analysis}
The latter part of the solution evaluation focuses on signal properties of vibrations from air compressors, water pumps, and electric induction motors. At first, the accelerometer data logger functionality is tested, and its precision is checked up in the known situations. Our custom dataset is compiled from collected measurements during regular operations of machines. The behaviour of different machines is compared using static frequency estimation and a time-frequency spectrogram. The methodology is carried out to diagnose the current status of water pumps. The presence of fault is confronted with sensor logs from the pump vendor.

@@

The electric motor signals are polluted by noise from the wind blowing out of a large cooling fan at the back of the motor. The M2 in place two compared to M1 has elevated amplitudes above 4 kHz. The pumps have richer signal content, likely due to the flow of water split into several frequency bands. The outer bearing (4) has a more attenuated amplitude above 1.5 kHz than the inner bearing (3). The P2 exhibits less vibration in comparable bands in general. The compressor casing produces a series of harmonics of rotational frequency, which are stronger near the scroll and suction valve than near the base.

@@

Domain experts recommended the procedure for fault identification by calculation of bearing characteristic frequencies (Tab.~\ref{tab:evaluation:bearing-freq-pump}), and then they approved the following results. Figure~\ref{fig:evaluation:bearing-freq} identifies harmonics of rotational speed and BPFO frequency in every machine and BPFI for M2-2. It can be assumed that those frequencies will be the reason for their damage in the future. The absolute acceleration is minuscule in current frequency bands and year-long sensor log of rms velocity (Fig~\ref{fig:evaluation:ksb-guard-rms-vibartions}).

@@

Therefore, the current results suggest that the bearings are in flawless condition. During the over five years the pump has been in service, there is not one instance of bearing fault due to rated lifespan and yearly prophylactic maintenance. This underscores the difficulty of the scarcity of non-faulty states in industrial environments. The monitoring is better suited for situations without preventive maintenance or with smaller machinery like the plunger pump BC~21/20S we considered during air conditioning inspections.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/pumps-TD-dim-3.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/pumps-FD-dim-3.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \caption{The atrribute value ranges of Pump dataset}
\end{figure}

The range of feature values for compressors and pumps are higher than in MaFaulDa despite most of them describing baseline status (Fig~\ref{fig:evaluation-sigma-pump-interference}). The increses can be seen for example on the interquartile range of rms (7.33 - 14.62 $\mathrm{m/s}^2$), centroid (3806 - 6003 Hz) or entropy (9.40 - 13.22 nats).

>>>>

\chapter{Evaluation} \label{chapter:evaluation}
The verification of proposed solutions for machinery fault diagnostics is focused on two activities, these are vibration measurement and defect identification. The accuracy of supervised learning using the k-nearest neighbour classifier is determined by the MaFaulDa dataset under various experiments. Data logger recording is compared to the known reference, and the collected Pump dataset is analyzed.

\section{Fault classifiction in MaFaulDa}
The four outlined experiments on MaFaulDa involve testing sets of features with very few members to see the effects in predictions when reducing the information content about machine status to the minimum. First, the attributes are left in the complete count after extraction. Then, all of their combinations are enumerated, and the resulting accuracy distribution is matched against accuracy after feature selection with similarity metrics.

@@

The label ``normal'' is not falsely attributed to other classes in either feature set, but other classes can get assigned to be ``normal''. Most mistakes happen while predicting outer race fault, which is confused with cage fault, imbalance, and less often with misalignment. The shaft imbalance is used for simulating bearing faults, which is a natural reason for this high error rate.

@@

Under every circumstance, the magnitude of the triaxial feature vector reaches better accuracy than those from only the axis of motion for the same source domain and bearing. The model for inner bearing A is more accurate than outer bearing B. The TD set is generally better in predictions than the FD set for the equivalent k-value. The relabeled dataset for high severity has a steeper decrease in accuracy for the same number of neighbours.

\subsection{Feature subset combinations}
The complete sets of predictors are even greatly shrunken to representation that could be presented in 3D plot or in perpendicular cross sections. These trend variables could be used in distinguishing faults the same way as rms amplitude indicates their presence. Each possible combinations of pairs, triplets, and quadruplets constructs a separate k-NN model on which prediction accuracy is evaluated.

@@

The decrease in accuracy with additional neighbours is apparent and similar to the trend in complete sets of features. We are interested in comparing maximum accuracies of the accuracy distribution because the optimal feature selection method tries to approach them. The reduction in maximal accuracy is more noticeable between three and five features 3 - 5\%, and almost the same amount between five and eleven neighbours.

@@

Table~\ref{tab:evaluation:fsel} compares the concrete case of choosing the three features from each domain on bearing A and k-NN with five neighbours. Table~\ref{tab:evaluation:fsel-severity} uses labels for high-severity faults. There is a significant difference in train and test accuracy, which means the model is likely overfitting. The percentile within the distribution is measured to its respective data, the train or validation set. However, the percentile of best features in the test data is calculated against train distribution.

Combining the rankings from several metrics is necessary to get consistent results. This is evidenced by the variability of the success of selected feature sets in final prediction performance under multiple conditions. The PCA with 3 components transformed from the complete feature sets is comparable in accuracy to selection methods with original attributes.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-TD-train.png}
        \caption{Time-domain features (train)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-FD-train.png}
        \caption{Frequency-domain features (train)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-TD-test.png}
        \caption{Time-domain features (test)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-FD-test.png}
        \caption{Frequency-domain features (test)}
    \end{subfigure}
    \caption{Model accuracy statistical distributions with feature selection methods for three predictors (k = 5)}
    \label{fig:evaluation:fsel-model-distr}
\end{figure}

@@

In 216 scenarios, the mutual information has better median accuracy (80.87\%) and distribution percentile (91.81\%) followed by rank product with an accuracy of 79.82\% of and percentile of 88.97\% (Tab.~\ref{tab:evaluation:compare-fsel-accuracy}). The scenarios are composed of 24 base dataset modifications and options for hyperparameters k-value and number of features.

The rank product is the best method in the majority of 43.52\% scenarios. Mutual information comes second in 40.28\% cases, where it is deemed the best strategy (Tab.~\ref{tab:evaluation:best-selection-method}). The median accuracy, in cases where they are the best method, is also better for rank product 92.38\% compared to 91.79\% mutual information.

@@

\subsection{Incremental learning}
Online learning imitates hardened conditions for machinery diagnostics that appear in deployment. Delayed provision or omission of actual labels undoubtedly degrades the reliability of the classification. The question is how quickly the accuracy approaches the optimal one from the nearest neighbors trained in batch and what is the effect on the classifier from routine difficulties associated with the continuous labeling process.

@@

The k-NN models in incremental learning experiments learn on the same base training dataset as in batch learning for bearing A. Online learning metrics are evaluated by progressive valuation on a dataset that is left unbalanced. In this manner, we can compare the training accuracies in the last sample of online models to their batch counterparts.

@@

A significant change in the data stream occurs after 294 out of 1438 observations (or after 737 for high severity faults) when all 49 (or 737) normal conditions are consumed in the training process. Counters of other faults show that predictions are skewed towards more represented imbalance and misalignment. The uneven evolution of categories in a stream impacts the development of accuracy in the remaining experiments. The test accuracies of comparable batch models from the three best features are 85.47\% (TD), 87.52\% (FD), 91.71\% (TD severity), and 91.94\% (FD severity).

@@

\section{Industrial dataset analysis}
The latter part of the solution evaluation focuses on signal properties of vibrations from air compressors, water pumps, and electric induction motors. At first, the accelerometer data logger functionality is tested, and its precision is checked up in the known situations. Our custom dataset is compiled from collected measurements during regular operations of machines. The behaviour of different machines is compared using static frequency estimation and a time-frequency spectrogram. The domain expert methodology is carried out to diagnose the current status of water pumps. The presence of fault is confronted with sensor logs from the pump vendor.

@@

The electric motor signals are polluted by noise from the wind blowing out of a large cooling fan at the back of the motor. The M2 in place two compared to M1 has elevated amplitudes above 4 kHz. The pumps have richer signal content than motors split into several frequency bands, likely due to the flow of water. The outer bearing (4) has a more attenuated amplitude above 1.5 kHz than the inner bearing (3). The P2 exhibits less vibration in comparable bands in general. The compressor casing produces a series of harmonics of rotational frequency, which are stronger near the scroll and suction valve than near the base.

@@

Domain experts recommended the procedure for fault identification by calculation of bearing characteristic frequencies (Tab.~\ref{tab:evaluation:bearing-freq-pump}), and they approved the following results. Figure~\ref{fig:evaluation:bearing-freq} identifies harmonics of rotational speed and BPFO frequency in every machine and BPFI for M2-2. It can be assumed that those frequencies will be the reason for their damage in the future. The absolute acceleration is minuscule in current frequency bands and year-long sensor log of rms velocity (Fig~\ref{fig:evaluation:ksb-guard-rms-vibartions}).

@@

Therefore, the current results suggest that the bearings are in flawless condition. During the over five years the pump has been in service, there is not one instance of bearing fault due to rated lifespan and yearly prophylactic maintenance. This underscores the scarcity of recording faulty states in industrial environments. The monitoring is better suited for situations without preventive maintenance or with smaller machinery like the plunger pump BC~21/20S we considered during air conditioning inspections.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/pumps-TD-dim-3.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/pumps-FD-dim-3.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \caption{The attribute value ranges of Pump dataset}
\end{figure}

The range of feature values for compressors and pumps are higher than in MaFaulDa despite most of them describing baseline status (Fig~\ref{fig:evaluation-sigma-pump-interference}). The increases can be seen for example on the interquartile range of rms (7.33 - 14.62 $\mathrm{m/s}^2$), centroid (3806 - 6003 Hz) or entropy (9.40 - 13.22 nats).
